---
categories:  
- ""    #the front matter should be like the one found in, e.g., blog2.md. It cannot be like the normal Rmd we used
- ""
date: "2021-09-30"
description: Risk-Return of DJIA stocks # the title that will show up once someone gets to this page
draft: false
image:  # save picture in \static\img\blogs. Acceptable formats= jpg, jpeg, or png . Your iPhone pics wont work

keywords: ""
slug: country
# slug is the shorthand URL address... no spaces plz
title: Gapminder country compare analysis
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
library(janitor)
```

# `gapminder` country comparison

You have seen the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. To get a glimpse of the dataframe, namely to see the variable names, variable types, etc., we use the `glimpse` function. We also want to have a look at the first 20 rows of data.

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

Your task is to produce two graphs of how life expectancy has changed over the years for the `country` and the `continent` you come from.

I have created the `country_data` and `continent_data` with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "China") # just choosing Greece, as this is where I come from

continent_data <- gapminder %>% 
            filter(continent == "Asia")
```

First, create a plot of life expectancy over time for the single country you chose. Map `year` on the x-axis, and `lifeExp` on the y-axis. You should also use `geom_point()` to see the actual data points and `geom_smooth(se = FALSE)` to plot the underlying trendlines. You need to remove the comments **\#** from the lines below for your code to run.

```{r, lifeExp_one_country}
 plot1 <- ggplot(country_data, mapping = aes(x = year , y = lifeExp))+
   geom_point() +
   geom_smooth(se = FALSE)


#Next we need to add a title. Create a new plot, or extend plot1, using the `labs()` function to add an informative title to the plot.


 plot1<- plot1 +
   labs(title = " Life Expectancy Over Time in China",
       x = "Year ",
       y = "Life Expectancy ") 


# plot1
```

Secondly, produce a plot for all countries in the *continent* you come from. (Hint: map the `country` variable to the colour aesthetic. You also want to map `country` to the `group` aesthetic, so all points for each country are grouped together).

```{r lifeExp_one_continent}
 ggplot(continent_data, mapping = aes(x = country  , y = lifeExp  , colour= country , group = country))+
   geom_point() + 
   geom_smooth(se = FALSE) 
```

Finally, using the original `gapminder` data, produce a life expectancy over time graph, grouped (or faceted) by continent. We will remove all legends, adding the `theme(legend.position="none")` in the end of our ggplot.

```{r lifeExp_facet_by_continent}
ggplot(data = gapminder , mapping = aes(x = year  , y = lifeExp  , colour= continent ))+
   geom_point() + 
   geom_smooth(se = FALSE) +
   facet_wrap(~continent) +
   theme(legend.position="none") 
```

Given these trends, what can you say about life expectancy since 1952? Again, don't just say what's happening in the graph. Tell some sort of story and speculate about the differences in the patterns.



Generally speaking, life expectancy across all continents have seen in increase since 1952, though certain countries have had small setbacks in the process. For example, China's life expectancy had a fall from 1957 to 1967. I think that was because of the "Great Leap Forward Movement", which caused the Great Chinese Famine and drastically decreased China's population and life expectancy.